BIP-Predictor

### 2025-08-08 Archive IWH Simon paper/09.08.2025/paper.tex
up

\subsection{MSSA Optimization Criterion}

Let $\mathbf{X}_t$ (of dimension $t\times n$) denote a set of $n$ explanatory series $\mathbf{x}_{1t},...,\mathbf{x}_{nt}$, with observations $\mathbf{x}_{it}=(x_{i1},...,x_{it})'$. Let $\mathbf{z}_t=(z_{1},...,z_t)$  denote a target series, which typically lies outside the linear space spanned by $\mathbf{X}_t$  (and is unknown at time $t$). For simplicity, we assume stationarity of all time series involved. In this framework, $\mathbf{X}_t$ corresponds to a matrix of selected economic indicators with $n=5$, and $z_{it}$, i=1,...,5,   is the output of a two-sided HP(160) filter applied to the four indicators as well as to GDP. Since $z_{it}$  depends on future observations $x_{it+1},x_{it+2},...$, the prediction task involves `tracking' $z_{it}$  using the estimate $y_{it}=\sum_{j=1}^ny_{ijt}$, where $y_{ijt}=\mathbf{b}_{ij}~'\mathbf{x}_{jt}=\sum_{k=0}^{L} b_{ijk}x_{j,t-k}$ are the outputs of a multivariate (causal or one-sided) filter $\mathbf{B}_i^f$, with columns $\mathbf{b}_{ij}$ of fixed length $L$, $j=1,...,5$, assigning weights to the last $L$ observations of the $j$-th indicator $\mathbf{x}_{jt}$ (the fixed-length assumption is merely used for ease of exposition).
For the sake of clarity, we now omit the index $i$ for all variables, assuming $z_t=z_{i_0t}$ for some fixed $i_0$.% Additionally, we assume a fixed filter length $L$, such that $\mathbf{B}^i_t=\mathbf{B}_t=\mathbf{B}$, which has dimension $L\times 5$.

@@

Unfortunately, classic predictors often generate excessively many false (`noisy') sign changes due to high-frequency leakage, as discussed above (see Figure \ref{hp_160}.)
%as illustrated by the amplitude function in Figure \eqref{hp_160} (bottom right panel) and further elaborated upon by \cite{Wildi2025}. 
 \cite{Wildi2025} shows that the SSA criterion helps control this phenomenon (see Fig.\ref{mssa_msse_zc} for illustration.) Using the dual formulation of the criterion \eqref{critssaht}, the author demonstrates that the SSA solution yields the lowest zero-crossing rate among all (linear) predictors with equivalent target correlation.
The concept can be extended to a multivariate framework, denoted M-SSA, by generalizing the correlation functions as detailed hereafter.

\subsection{Implementation}

For a better understanding, we here briefly substantiate the optimization criterion \ref{critssa}. To start, assume a simple univariate framework with $x_t=\epsilon_t$ white noise and $z_t=\sum_{k=-infty}^\infty \gamma_k x_{t-k}=\sum_{k=-infty}^\infty \gamma_k \epsilon_{t-k}$, where it is assumed that $\gamma_k\neq 0$ for some $k<0$ (acausal filter) and $\sum_{|k|<\infty}\gamma_k^2<\infty$, which ensures stationarity of $z_t$ (extensions to non-stationary integrated processes are discussed in Wildi (2025)). Assume we want to predict $z_{t+\delta}$, $\delta\geq 0$ and let $y_{\delta t}=\sum_{k=0}^{L-1} b_{\delta k}x_{t-k}$ designate a causal predictor of length $L$. For simplicity of exposition we may assume $\delta=0$ (nowcast) and we may drop $\delta$ from our notation. Then, $\rho(z,y)=\frac{\sum_{k=0}^{L-1}b_k\gamma_k}{\sqrt{\sum_{k=0}^{L-1}b_k^2}\sqrt{\sum_{k=0}^{L-1}\gamma_k^2}}$  and $\rho(y,1)=\frac{\mathbf{b}'\mathbf{Mb}}{\mathbf{b}'\mathbf{b}}$, where 
\[
\mathbf{M}=\left(\begin{array}{ccccccccc}0&0.5&0&0&0&...&0&0&0\\
0.5&0&0.5&0&0&...&0&0&0\\
...&&&&&&&&\\
0&0&0&0&0&...&0.5&0&0.5\\
0&0&0&0&0&...&0&0.5&0
\end{array}\right),
\]
ofs dimension $L\times L$ is such that $\mathbf{b}'\mathbf{Mb}=\sum_{k=1}^{L-1}b_{k-1}b_k$ corresponds to the order-one autocovariance of $y_t$, up to scaling by the variance $\sigma^2$ of $\epsilon_t$. We can now rewrite criterion \ref{critssa} as:
\begin{eqnarray}\label{crit1}
\left.\begin{array}{cc}
&\max_{\mathbf{b}}\mathbf{b}'\boldsymbol{\gamma}_{\delta}\\
&\mathbf{b}'\mathbf{Mb}=l\rho_1\\
&\mathbf{b}'\mathbf{b}=l
\end{array}\right\}.
\end{eqnarray}
where the additional length constraint $\mathbf{b}'\mathbf{b}=l$ ensures uniqueness of the SSA solution. 
By orthogonal projection, the MSE predictor is $y_{MSE,t}=\sum_{k=0}^{L-1}\gamma_kx_{t-k}$, i.e., $b_{MSE,k}=\gamma_k$, $k=0,...,L-1$ and in general we select $\rho_1>\rho_{MSE}$ in criterion l{crit1} to improve smoothness of the resulting SSA predictor $y_t$, where $\rho_{MSE}$ designates the first order ACF of the (benchmark) MSE predictor\footnote{Selecting $\rho_1<\rho_{MSE}$ would result in a SSA predictor with more frequent zero-crossings than the MSE benchmark, which is undesirable in typical applications.}. \\

We can extend the SSA criterion \ref{crit1} to a multivariate framework. Assume first that $\mathbf{X}_t=\boldsymbol{\epsilon}_t$ is a multivariate white noise sequence with (full rank) variance-covariance matrix $\boldsymbol{\Sigma}$ and let $\mathbf{z}_t$:

@@

where $\boldsymbol{\Gamma}_k$ is a sequence of square-summable $n\times n$ dimensional filter matrices, with entries $(\gamma_{ijk})$ for $i, j = 1, \ldots, n$. The condition for square summability is expressed mathematically as:
\sum_{|k| < \infty} \text{tr}(\boldsymbol{\Gamma}_k' \boldsymbol{\Gamma}_k) < \infty,
where $\text{tr}(\cdot)$ denotes the trace operator. We focus on the estimation of $\mathbf{z}_{t+\delta}$, utilizing the $n-$dimensional predictor defined by

@@

where $\boldsymbol{\gamma}_{i k}:=(\gamma_{i1k},...,\gamma_{ink})'$. Consequntly, we propose the following multivariate M-SSA criterion for $i=1,...,n$: 

@@

\mathbf{b}_{i}'\tilde{\mathbf{M}}\mathbf{b}_{i}=\rho_i\nonumber\\

@@

Under the imposed length constraint, the objective function is proportional to the correlation between the predictor and the $i$-th target, or equivalently, the $i$-th MSE predictor. Additionally, the HT constraint regulates the first-order ACF or, equivalently, the HT of the predictor. The relationships established between sign accuracy and target correlation, as well as between first-order ACF and HT, as discussed in the previous section, remain valid. Thus, the M-SSA framework can be reformulated in accordance with any alternative and equivalent expressions of the optimization criterion, as reviewed in the previous section. \\

To conclude, we now relax the WN hypothesis and assume a stationary process 

@@

(\mathbf{b}\cdot\boldsymbol{\xi})_i'\mathbf{\tilde{M}}(\mathbf{b}\cdot\boldsymbol{\xi})_i=\rho_i\nonumber\\

@@

for $i=1,...,n$. This formulation can be solved for $(\mathbf{b}\cdot\boldsymbol{\xi})_i$, as detailed in Theorem \eqref{lambda_mult} in the next section. The M-SSA coefficients $\mathbf{b}_i$ can subsequently be determined through deconvolution, as indicated in Equation \eqref{con_inv}. 
Wildi (2025) posits an extension of SSA to non-stationary integrated processes, wherein the HT constraint pertains to stationary differences of the predictor. While a corresponding extension to M-SSA is feasible based on findings by McElroy and Wildi (2020), the present discussion is confined to stationary processes, omitting a lengthier treatment of rank-defficient designs (cointegration) in the non-stationary case.\\

%\textcolor{blue}{This is where Marc outlines the generalization of the correlation function and other elements of M-SSA.}

@@

%Lastly, sign changes in trend growth, represented by zero-crossings of the target $z_t$, are indicative of relevant changes in the trajectory of the economy, for example, at transitions between expansions and recessions.

 %In a second stage, serving as a generalization and refinement of the univariate HP-C utilized in the direct HP forecast.

@@

\begin{figure}[H]
    \begin{center}
    \textcolor{blue}{@Marc: Please change the plot Title to "M-MSE \& M-SSA Nowcasts: Financial Crisis".)}
        \includegraphics[height=3in, width=4.5in]{./Figures/mssa_msse_zc.pdf}
        % \caption{Sign changes of the M-MSE (green) and M-SSA (blue) nowcasts towards the financial crisis, with sign changes of the M-MSE marked by green vertical lines. The sample holding time (HT) of each filter corresponds to the average duration between its consecutive zero-crossings. For each filter, its sample HT converges to the respective expected HT asymptotically. Here, we impose  $HT_{MSSA}=1.5\times HT_{MSE}$: in the long run, we expect the M-MSE to exhibit approximately $50\%$ more sign changes than the M-SSA.
        \caption{M-MSE (green) and M-SSA (blue) nowcasts around the financial crisis.\\
        Sign changes of the M-MSE marked by green vertical lines.\\
        The holding time (HT) of each filter corresponds to the expected duration between its sign-changes. 
        % For each filter, its sample HT converges to the respective expected HT asymptotically. 
        Here, we impose  $HT_{MSSA}=1.5\times HT_{MSE}$. In large samples, we expect the M-SSA to exhibit approximately $33\%$ fewer sign changes than the M-MSE.
        \label{mssa_msse_zc}}
    \end{center}
\end{figure}
% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Wed May  7 12:49:47 2025
\begin{table}[ht]
\centering
\begin{tabular}{lrr}
  \hline
 & M-SSA & M-MSE \\ 
  \hline
Target correlation & 0.68 & 0.69 \\ 
  HT & 6.50 & 3.71 \\ 
   \hline
\end{tabular}
\caption{Sample target correlations and holding times (HT) of M-SSA and M-MSE nowcasts.} 
\label{corhtnow}
\end{table}

>>>>

\subsection{SSA Optimization Criterion}

Let $\mathbf{x}_t$ be a time series representing the data, with observations $(x_{1},...,x_{t})'$. Let $\mathbf{z}_t=(z_{1},...,z_t)$  denote a target series, which depends on $\mathbf{x}_t$ as well as on unknown future $x_{t+k}$, $k>0$. For simplicity, we assume stationarity of all time series involved. In this framework, we might think of ${z}_t$ as corresponding to the output of a two-sided HP(160) filter applied to the doubly-infinite sequence $\mathbf{x}_{\pm \infty}=(x_k)$, $|k|<\infty$, representing the (log differenced) GDP, for which $\mathbf{x}_t$ is effectively observed. Since $z_{t}$  depends on future observations $x_{t+1},x_{t+2},...$\footnote{We here ignore the dependence of $z_t$ on $x_0,x_{-1},...$ since in general the filter weights $\gamma_k$ of the target decay sufficiently fast to zero.}, the prediction task involves `tracking' $z_{t}$  using the (causal) predictor $y_{t}=\sum_{k=0}^{L-1} b_{jk}x_{t-k}$, where $L\leq t$.

@@

Unfortunately, classic predictors often generate excessively many false (`noisy') sign changes due to high-frequency leakage, as discussed above (see Figure \ref{hp_160}.)
%as illustrated by the amplitude function in Figure \eqref{hp_160} (bottom right panel) and further elaborated upon by \cite{Wildi2025}. 
 \cite{Wildi2025} shows that the SSA criterion helps control this phenomenon (see Fig.\ref{mssa_msse_zc} for illustration.) Using the dual formulation of the criterion \eqref{critssaht}, the author demonstrates that the SSA solution yields the lowest zero-crossing rate among all (linear) predictors with equivalent target correlation.
The concept can be extended to a multivariate framework, denoted M-SSA, by generalizing the correlation functions as detailed in the next section.\\

We now briefly elucidate the SSA optimization criterion outlined in Equation \ref{critssa}. Consider a simplified univariate framework where $x_t=\epsilon_t$ represents white noise with variance $\sigma^2$. A generic target specification for $z_t$ is then defined as a linear filter of $x_t$:
\[z_t=\sum_{k=-\infty}^\infty \gamma_k x_{t-k}=\sum_{k=-\infty}^\infty \gamma_k \epsilon_{t-k},\] 
where it is assumed that $\gamma_k\neq 0$ for some $k<0$ (implying an acausal filter), and the summation of squared coefficients satisfies $\sum_{|k|<\infty}\gamma_k^2<\infty$, ensuring the stationarity of the target $z_t$. With these definitions, all processes are centered at zero and the Gaussian assumption additionally justifies the use of Equation \ref{critssaht}, linking first-order ACF and HT. Extensions to non-stationary integrated processes are discussed in Wildi (2025). Suppose the goal is to predict  $z_{t+\delta}$, $\delta\geq 0$. Let  $y_{\delta t}=\sum_{k=0}^{L-1} b_{\delta k}x_{t-k}$  denote a causal predictor of length $L$.  To simplify the exposition, we consider the case $\delta=0$ (nowcast), and thus omit $\delta$ from the notation. Then, under the postulated white assumption
\begin{eqnarray*}
\rho(z,y)&=&\frac{\sum_{k=0}^{L-1}b_k\gamma_k}{\sqrt{\sum_{k=0}^{L-1}b_k^2}\sqrt{\sum_{|k|<\infty}\gamma_k^2}},\\
\rho(y,1)&=&\frac{\mathbf{b}'\mathbf{Mb}}{\mathbf{b}'\mathbf{b}},
\end{eqnarray*} 
where the so-called autocovariance generating matrix $\mathbf{M}$, of dimension $L\times L$, is defined as:
\[
\mathbf{M}=\left(\begin{array}{ccccccccc}0&0.5&0&0&0&...&0&0&0\\
0.5&0&0.5&0&0&...&0&0&0\\
...&&&&&&&&\\
0&0&0&0&0&...&0.5&0&0.5\\
0&0&0&0&0&...&0&0.5&0
\end{array}\right).
\]
The matrix $\mathbf{M}$ is such that $\mathbf{b}'\mathbf{Mb}=\sum_{k=1}^{L-1}b_{k-1}b_k$, representing the first-order autocovariance of $y_t$ (up to the scaling factor $\sigma^2$).  Using this, the optimization criterion \ref{critssa} can be reformulated as:
\begin{eqnarray}\label{crit1}
\left.\begin{array}{cc}
&\max_{\mathbf{b}}\mathbf{b}'\boldsymbol{\gamma}_{\delta}\\
\textrm{subject~to~}&\mathbf{b}'\mathbf{Mb}=l\rho_1\\
&\mathbf{b}'\mathbf{b}=l
\end{array}\right\},
\end{eqnarray}
where the additional length constraint $\mathbf{b}'\mathbf{b}=l$ serves two purposes: firstly, to ensure the uniqueness of the SSA solution; secondly, to normalize the objective function $\mathbf{b}'\boldsymbol{\gamma}_{\delta}$, making it proportional to the target correlation, up to a fixed scaling factor $\sqrt{\mathbf{b}'\mathbf{b}}\sqrt{\sum \gamma_k^2}=\sqrt{l}\sqrt{\sum \gamma_k^2}$. Criterion \ref{crit1} is readily amenable to optimization, as it involves affine and quadratic functions of $\mathbf{b}$, see Wildi (2025) for detailed derivations. \\
Under the above assumptions, the classical MSE predictor can be expressed via orthogonal projection as:
\[
y_{MSE,t}=\sum_{k=0}^{L-1}\gamma_kx_{t-k}
\]
which corresponds to setting $b_{MSE,k}=\gamma_k$, $k=0,...,L-1$. In general, to enhance the smoothness of the SSA predictor $y_t$, we select $\rho_1>\rho_{MSE}$ in criterion \ref{crit1},  where $\rho_{MSE}$ denotes the first-order ACF of the (benchmark) MSE predictor\footnote{Selecting $\rho_1<\rho_{MSE}$ would produce an SSA predictor with more frequent zero-crossings than the MSE benchmark, which is typically undesirable in practical applications.}.

\subsection{M-SSA Optimization}

We can now extend the SSA criterion \ref{crit1} to a multivariate setting. Let $\mathbf{X}_t$ (of dimension $t\times n$) denote a set of $n$ explanatory series $\mathbf{x}_{1t},...,\mathbf{x}_{nt}$, with observations $\mathbf{x}_{it}=(x_{i1},...,x_{it})'$. Let $\mathbf{z}_t=(z_{1},...,z_t)$  denote a target series, which typically lies outside the linear space spanned by $\mathbf{X}_t$  (and is unknown at time $t$). In this framework, $\mathbf{X}_t$ corresponds to a matrix of selected economic indicators with $n=5$, and $z_{it}$, i=1,...,5,   is the output of a two-sided HP(160) filter applied to the four indicators as well as to GDP. %Since $z_{it}$  depends on future observations $x_{it+1},x_{it+2},...$, the prediction task involves `tracking' $z_{it}$  using the estimate $y_{it}=\sum_{j=1}^ny_{ijt}$, where $y_{ijt}=\mathbf{b}_{ij}~'\mathbf{x}_{jt}=\sum_{k=0}^{L} b_{ijk}x_{j,t-k}$ are the outputs of a multivariate (causal or one-sided) filter $\mathbf{B}_i^f$, with columns $\mathbf{b}_{ij}$ of fixed length $L$, $j=1,...,5$, assigning weights to the last $L$ observations of the $j$-th indicator $\mathbf{x}_{jt}$ (the fixed-length assumption is merely used for ease of exposition).
For the sake of clarity, we now omit the index $i$ for all variables, assuming $z_t=z_{i_0t}$ for some fixed $i_0$.

To introduce the M-SSA, assume that $\mathbf{X}_t=\boldsymbol{\epsilon}_t$ is an $n$-dim white noise sequence with (full rank) variance-covariance matrix $\boldsymbol{\Sigma}$. Consider the generalized target $\mathbf{z}_t$:

@@

where $\boldsymbol{\Gamma}_k$  is a sequence of filter matrices of dimension  $n\times n$, satisfying the square-summability condition. Each $\boldsymbol{\Gamma}_k$ has entries $(\gamma_{ijk})$ for $i, j = 1, \ldots, n$. We focus on the estimation of $\mathbf{z}_{t+\delta}$, utilizing the $n-$dimensional predictor defined by

@@

where $\boldsymbol{\gamma}_{i k}:=(\gamma_{i1k},...,\gamma_{ink})'$. Consequently, we propose the following multivariate M-SSA criterion for $i=1,...,n$: 

@@

\textrm{subject~to~}\mathbf{b}_{i}'\tilde{\mathbf{M}}\mathbf{b}_{i}=\rho_i\nonumber\\

@@

Under the imposed length constraint, the objective function is proportional to the correlation between the predictor and the $i$-th target. Furthermore, the criterion relies on affine and quadratic functions of $\mathbf{b}_i$, facilitating the numerical optimization. Finally, the HT constraint regulates the first-order ACF or, equivalently, the HT of the predictor.\\

Finally, we can relax the WN hypothesis and assume a stationary process 

@@

\textrm{subject~to~}(\mathbf{b}\cdot\boldsymbol{\xi})_i'\mathbf{\tilde{M}}(\mathbf{b}\cdot\boldsymbol{\xi})_i=\rho_i\nonumber\\

@@

for $i=1,...,n$. In analogy to the above white noise case, this formulation can be solved for $(\mathbf{b}\cdot\boldsymbol{\xi})_i$. The M-SSA coefficients $\mathbf{b}_i$ can subsequently be determined through deconvolution, as indicated in Equation \eqref{con_inv}. 
The present discussion is confined to stationary processes, omitting a lengthier treatment of rank-defficient designs (cointegration) in the non-stationary case.\\

@@

@@

\begin{figure}[H]
    \begin{center}
%    \textcolor{blue}{@Marc: Please change the plot Title to "M-MSE \& M-SSA Nowcasts: Financial Crisis".)}
        \includegraphics[height=3in, width=4.5in]{./Figures/mssa_msse_zc.pdf}
        % \caption{Sign changes of the M-MSE (green) and M-SSA (blue) nowcasts towards the financial crisis, with sign changes of the M-MSE marked by green vertical lines. The sample holding time (HT) of each filter corresponds to the average duration between its consecutive zero-crossings. For each filter, its sample HT converges to the respective expected HT asymptotically. Here, we impose  $HT_{MSSA}=1.5\times HT_{MSE}$: in the long run, we expect the M-MSE to exhibit approximately $50\%$ more sign changes than the M-SSA.
        \caption{M-MSE (green) and M-SSA (blue) nowcasts around the financial crisis.\\
        Sign changes of the M-MSE marked by green vertical lines.\\
        The holding time (HT) of each filter corresponds to the expected duration between its sign-changes. 
        % For each filter, its sample HT converges to the respective expected HT asymptotically. 
        Here, we impose  $HT_{MSSA}=1.5\times HT_{MSE}$. In large samples, we expect the M-SSA to exhibit approximately $33\%$ fewer sign changes than the M-MSE.
        \label{mssa_msse_zc}}
    \end{center}
\end{figure}
% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Wed May  7 12:49:47 2025
\begin{table}[ht]
\centering
\begin{tabular}{lrr}
  \hline
 & M-SSA & M-MSE \\ 
  \hline
Target correlation & 0.68 & 0.69 \\ 
  HT & 6.50 & 3.71 \\ 
   \hline
\end{tabular}
\caption{Sample target correlations and holding times (HT) of M-SSA and M-MSE nowcasts.} 
\label{corhtnow}
\end{table}

### 2025-08-08 Sweave/BIP_predictor.Rnw
up

# We can select type of VAR model
VAR_type="BVAR"
VAR_type="VAR"
VAR_type="elastic_net"
if (VAR_type=="BVAR")
{
# Simon's settings for BVAR(4)  
  lambda_BVAR <- 1.001
  p<-3
  q<-0
}
if (VAR_type=="VAR")
{
# Marc's settings for VAR  
  p<-1
  q<-0
}
if (VAR_type=="elastic_net")
{
# Marc's settings for EN
  q<-0
# 1. Simpler (but possibly too simple)  
  p<-2
  nsteps<-5
# 2. Simpler than Simon's original settings (below) but slightly more complex than above ones  (good compromise)
  p<-3
  nsteps<-6
# Simon's original settings (lead to overfitting)  
#  p<-4
#  nsteps<-7
}

@@

colo<-c("black","violet","blue","orange","green")
colo<-c("blue","green")
main_title<-"Zero-Crossings of M-SSA and M-MSE towards the financial crisis"
plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],lwd=1,ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
mtext(colnames(mplot)[1],col=colo[1],line=-1)
for (i in 1:ncol(mplot))
{
  lines(mplot[,i],col=colo[i],lwd=1,lty=1)
  mtext(colnames(mplot)[i],col=colo[i],line=-i)
}
abline(h=0)
abline(v=1+which(sign(mplot[2:nrow(mplot),1])!=sign(mplot[1:(nrow(mplot)-1),1])),col=colo[1],lty=2)
abline(v=1+which(sign(mplot[2:nrow(mplot),2])!=sign(mplot[1:(nrow(mplot)-1),2])),col=colo[2],lty=2)
axis(1,at=c(1,4*1:(nrow(mplot)/4)),labels=rownames(mplot)[c(1,4*1:(nrow(mplot)/4))])
axis(2)
box()

>>>>

# We can select type of VAR model
#VAR_type="BVAR"
VAR_type="VAR"
#VAR_type="elastic_net"
if (VAR_type=="BVAR")
{
# Simon's settings for BVAR(4)  
  lambda_BVAR <- 1.001
  p<-3
  q<-0
}
if (VAR_type=="VAR")
{
# Marc's settings for VAR  
  p<-1
  q<-0
}
if (VAR_type=="elastic_net")
{
# Marc's settings for EN
  q<-0
# 1. Simpler (but possibly too simple)  
  p<-2
  nsteps<-5
# 2. Simpler than Simon's original settings (below) but slightly more complex than above ones  (good compromise)
  p<-3
  nsteps<-6
# Simon's original settings (lead to overfitting)  
#  p<-4
#  nsteps<-7
}

@@

colo<-c("black","violet","blue","orange","green")
colo<-c("blue","green")
main_title<-"M-MSE & M-SSA Nowcasts: Financial Crisis"
plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],lwd=1,ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
mtext(colnames(mplot)[1],col=colo[1],line=-1)
for (i in 1:ncol(mplot))
{
  lines(mplot[,i],col=colo[i],lwd=1,lty=1)
  mtext(colnames(mplot)[i],col=colo[i],line=-i)
}
abline(h=0)
abline(v=1+which(sign(mplot[2:nrow(mplot),1])!=sign(mplot[1:(nrow(mplot)-1),1])),col=colo[1],lty=2)
abline(v=1+which(sign(mplot[2:nrow(mplot),2])!=sign(mplot[1:(nrow(mplot)-1),2])),col=colo[2],lty=2)
axis(1,at=c(1,4*1:(nrow(mplot)/4)),labels=rownames(mplot)[c(1,4*1:(nrow(mplot)/4))])
axis(2)
box()
